---
title: "Modeling Demand"
author: "Gustavo Mello"
date: '2022-03-28'
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(tsibble)
library(fable)
library(feasts)
library(bizdays)
library(ggdark)

theme_set(dark_theme_gray())
```

# Data loading
```{r loading}
# Imports dataset made in EDA script
cwd <- getwd()
daily.revenue.listings <- read_csv(paste0(cwd, "/../data/output/daily_revenue_listings.csv"))
```

# Reservations across time
```{r reservations}

reservations.daily <- daily.revenue.listings %>%
  filter(date <= "2022-03-15", occupancy==1, blocked==0) %>%
  select(date) %>%
  group_by(date) %>%
  summarise(count=n()) %>%
  tsibble(index=date) %>%
  fill_gaps(count=0, .full=TRUE) %>%
  mutate(is.workday = !(timeDate::isWeekend(date)
                        | (date %in% holidaysANBIMA)))

autoplot(reservations.daily)
```


## Forecasting
### Dynamic Harmonic Regression
To forecast with this dataset, we need a model that captures multiple seasonalities and is robust to sudden changes (like the pandemic). Also, as the seasonality effects seems to scale with the response variable magnitude, it is reasonable to test applying a log transform prior to modeling. For these reasons, a dynamic harmonic regression model is fitted using Fourier terms to account for seasonality and an ARMA model to account for short-term dynamics.
```{r dynamic_harmonic_regression}
dynamic_harmonic_regression <- function(K, log_transform=FALSE){
  K_weekly <- if (K > 3) {3} else {K}
  if(log_transform){
    ARIMA(log(count) ~ PDQ(0, 0, 0)
                       + fourier(K=K_weekly, period=7)
                       + fourier(K=K, period=365.25)
                       + is.workday)
  } else{
    ARIMA(count ~ PDQ(0, 0, 0)
                  + fourier(K=K_weekly, period=7)
                  + fourier(K=K, period=365.25)
                  + is.workday)
  }
}

fit <- reservations.daily %>%
  mutate(count=count + 1) %>%
  model("K = 1"      = dynamic_harmonic_regression(K=1),
        "K = 2"      = dynamic_harmonic_regression(K=2),
        "K = 3"      = dynamic_harmonic_regression(K=3),
        "K = 4"      = dynamic_harmonic_regression(K=4),
        "K = 5"      = dynamic_harmonic_regression(K=5),
        "K = 6"      = dynamic_harmonic_regression(K=6),
        "K = 1, log" = dynamic_harmonic_regression(K=1, log_transform = TRUE),
        "K = 2, log" = dynamic_harmonic_regression(K=2, log_transform = TRUE),
        "K = 3, log" = dynamic_harmonic_regression(K=3, log_transform = TRUE),
        "K = 4, log" = dynamic_harmonic_regression(K=4, log_transform = TRUE),
        "K = 5, log" = dynamic_harmonic_regression(K=5, log_transform = TRUE),
        "K = 6, log" = dynamic_harmonic_regression(K=6, log_transform = TRUE))
```

``` {r plot_forecasts, fig.width=12, fig.heigth=24}
reservations.daily.new.data <- reservations.daily %>%
  new_data(291L) %>%
  mutate(is.workday = !(timeDate::isWeekend(date)
                        | (date %in% holidaysANBIMA)))

fit %>%
  forecast(new_data=reservations.daily.new.data) %>%
  autoplot(reservations.daily, level=NULL) +
  facet_wrap(vars(.model), ncol=2, scale="free_y") +
  geom_label(
    aes(x = ymd("2021-01-01"), y = 200,
        label = paste0("AICc = ", format(AICc))),
    data = glance(fit))
```

### DHR excluding COVID lockdowns in 2022 and prior data
```{r exclude_covid}
reservations.daily.COVID.excluded <- reservations.daily %>%
  filter_index("2021-01-01" ~ "2022-03-15")

fit.COVID.excluded <- reservations.daily.COVID.excluded %>%
  mutate(count=count + 1) %>%
  model("K = 1"      = dynamic_harmonic_regression(K=1),
        "K = 2"      = dynamic_harmonic_regression(K=2),
        "K = 3"      = dynamic_harmonic_regression(K=3),
        "K = 4"      = dynamic_harmonic_regression(K=4),
        "K = 5"      = dynamic_harmonic_regression(K=5),
        "K = 6"      = dynamic_harmonic_regression(K=6),
        "K = 1, log" = dynamic_harmonic_regression(K=1, log_transform = TRUE),
        "K = 2, log" = dynamic_harmonic_regression(K=2, log_transform = TRUE),
        "K = 3, log" = dynamic_harmonic_regression(K=3, log_transform = TRUE),
        "K = 4, log" = dynamic_harmonic_regression(K=4, log_transform = TRUE),
        "K = 5, log" = dynamic_harmonic_regression(K=5, log_transform = TRUE),
        "K = 6, log" = dynamic_harmonic_regression(K=6, log_transform = TRUE))
```

``` {r plot_forecasts_COVID_excluded, fig.width=12, fig.heigth=24}
reservations.daily.COVID.excluded.new.data <- reservations.daily.COVID.excluded %>%
  new_data(291L) %>%
  mutate(is.workday = !(timeDate::isWeekend(date)
                        | (date %in% holidaysANBIMA)))

fit.COVID.excluded %>%
  forecast(new_data=reservations.daily.COVID.excluded.new.data) %>%
  autoplot(reservations.daily.COVID.excluded,
           level=NULL) +
  facet_wrap(vars(.model), ncol=2, scale="free_y") +
  geom_label(
    aes(x = ymd("2021-06-01"), y = 200,
        label = paste0("AICc = ", format(AICc))),
    data = glance(fit))
```

```{r plot_K}
plot_K <- function(K=2){
  fit_test <- reservations.daily.COVID.excluded %>%
    mutate(count=count + 1) %>%
    model(dynamic_harmonic_regression(K=K, log_transform = TRUE))
  
    fit_test %>%
  forecast(new_data=reservations.daily.COVID.excluded.new.data) %>%
  autoplot(reservations.daily.COVID.excluded.new.data,
           level=NULL) +
  geom_label(
    aes(x = ymd("2021-06-01"), y = 200,
        label = paste0("AICc = ", format(AICc))),
    data = glance(fit))
}
```

```{r cross_validating_models}
models <- reservations.daily.COVID.excluded %>%
  filter_index(~ "2022-03-04") %>%
  stretch_tsibble(.init = 369, .step=10) %>%
  model(
    "K = 1"  = dynamic_harmonic_regression(K=1,  log_transform=TRUE),
    "K = 2"  = dynamic_harmonic_regression(K=2,  log_transform=TRUE),
    "K = 5"  = dynamic_harmonic_regression(K=5,  log_transform=TRUE),
    "K = 10" = dynamic_harmonic_regression(K=10, log_transform=TRUE),
    "K = 20" = dynamic_harmonic_regression(K=20, log_transform=TRUE),
    "K = 30" = dynamic_harmonic_regression(K=30, log_transform=TRUE),
    "K = 50" = dynamic_harmonic_regression(K=50, log_transform=TRUE)
  )
```

```{r cross_validating_selection}
CV.data <- reservations.daily.COVID.excluded %>%
  filter_index(~ "2022-03-04") %>%
  stretch_tsibble(.init = 369, .step=10) %>%
  new_data(10L) %>%
  mutate(is.workday = !(timeDate::isWeekend(date)
                        | (date %in% holidaysANBIMA)))

results <- models %>%
  forecast(new_data=CV.data) %>%
  accuracy(reservations.daily.COVID.excluded) %>%
  select(.model, RMSE:MAPE)

results %>% arrange(RMSE)
```

```{r cross_validate_one}
cross.validate.for <- function(K){
  model_for_K <- reservations.daily.COVID.excluded %>%
    filter_index(~ "2022-03-04") %>%
    stretch_tsibble(.init = 369, .step=10) %>%
    model(dynamic_harmonic_regression(K=K,  log_transform=TRUE))
  
  CV.data <- reservations.daily.COVID.excluded %>%
  filter_index(~ "2022-03-04") %>%
  stretch_tsibble(.init = 369, .step=10) %>%
  new_data(10L) %>%
  mutate(is.workday = !(timeDate::isWeekend(date)
                        | (date %in% holidaysANBIMA)))

  model_for_K %>%
  forecast(new_data=CV.data) %>%
  accuracy(reservations.daily.COVID.excluded) %>%
  select(.model, RMSE:MAPE)
}
```

```{r optimal_parameter}
parameter_metrics <- results %>%
  mutate(K=sapply(results$.model,
                  function(x){
                    parse_integer(substring(x, first=5))
                  })) %>%
  add_row(.model="K = 100", RMSE=98.63, MAE=77.58,
          MPE=6.63, MAPE=38.10, K=100L) %>%
  add_row(.model="K = 100", RMSE=98.63, MAE=77.58,
          MPE=6.63, MAPE=38.10, K=100L) %>%
  add_row(.model="K = 60", RMSE=67.92, MAE=56.39,
          MPE=6.19, MAPE=27.67, K=60L) %>%
  add_row(.model="K = 40", RMSE=79.48, MAE=63.04,
          MPE=5.30, MAPE=30.54, K=40L) %>%
  add_row(.model="K = 70", RMSE=63.83, MAE=54.16,
          MPE=3.14, MAPE=26.76, K=70L) %>%
  add_row(.model="K = 80", RMSE=92.31, MAE=73.40,
          MPE=12.94, MAPE=35.53, K=80L) %>%
  add_row(.model="K = 65", RMSE=66.08, MAE=56.02,
          MPE=3.40, MAPE=27.61, K=65L) %>%
  add_row(.model="K = 75", RMSE=61.54, MAE=51.18,
          MPE=2.17, MAPE=25.34, K=75L) %>%
  add_row(.model="K = 77", RMSE=65.98, MAE=55.84,
          MPE=0.93, MAPE=27.39, K=77L) %>%
  add_row(.model="K = 73", RMSE=64.54, MAE=53.98,
          MPE=1.01, MAPE=26.69, K=73L) %>%
  arrange(K)

parameter_metrics

parameter_metrics %>%
  pivot_longer(RMSE:MAPE, names_to = "Metric", values_to = "Value") %>%
  ggplot(aes(x=K, y=Value, color=Metric)) +
  geom_point(size=2) +
  geom_line() +
  geom_smooth(se=FALSE, linetype="dashed", size=0.5)
```
```{r final_model}
final.model <- reservations.daily.COVID.excluded %>%
  model(dynamic_harmonic_regression(K=75, log_transform=TRUE))
```

```{r final_forecast}
final.model %>%
  forecast(reservations.daily.COVID.excluded.new.data) %>%
  autoplot(reservations.daily.COVID.excluded, level=80) +
  labs("Daily Number of Reservations Forecast",
       x="Day",
       y="Reservations")

final.model %>%
  forecast(reservations.daily.COVID.excluded.new.data) %>%
  autoplot(reservations.daily.COVID.excluded, level=80) +
  labs("Daily Number of Reservations Forecast",
       x="Day",
       y="Reservations (log scale)") +
  scale_y_log10()
```

